---
output: 
  github_document:
    html_preview: false
# params: 
#   actor_id: "esch" 
#   data_date: "2020-03-26"
#   sha: "sha"
---

### Buffer distance

-----

Buffer distances (in miles) extend from the perimeter of units. For trails, the buffer distance is effectively a radius (i.e. a 1.0 mile buffer on each side of a trail would mean a 2.0 mile diameter path centered on the trail.) <br><br>

```{r include=F}
library(tidyverse)
library(sf)

park <- regionalparks.acs::park_trail_geog_LONG[c(15),0] %>%
  st_transform(3857)
buf025 <- park  %>%
  st_buffer(dist = 1609.34 * .25)
buf05 <- park  %>%
  st_buffer(dist = 1609.34 * .5)

trail <- regionalparks.acs::park_trail_geog_LONG[c(109),0] %>%
  st_transform(3857)
tbuf025 <- trail  %>%
  st_buffer(dist = 1609.34 * .25)
tbuf05 <- trail  %>%
  st_buffer(dist = 1609.34 * .5)
```

Example of two different buffer distances (dark grey, light grey) around a park polygon (green)

```{r echo=F, fig.width=2, fig.height=2}
ggplot() +
  geom_sf(data = buf05, fill = "grey90") +
  geom_sf(data = buf025, fill = "grey70") +
  geom_sf(data = park, fill = "#78A22F") +
  theme_void()

```
 
<br>
Example of two different buffer distances (dark grey, light grey) around a trail segment (green)
<br>

```{r echo=F, message=F, warning=F, fig.width=2, fig.height=2}
ggplot() +
  geom_sf(data = tbuf05, fill = "grey90") +
  geom_sf(data = tbuf025, fill = "grey70") +
  geom_sf(data = trail, color = "#78A22F", size=4) +
  theme_void()
```
 


